---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2021"
subtitle: "Assignment 4 - Due date 02/25/21"
author: "Xueying Feng"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: inline
---
  
## Directions
  
  You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. And to do so you will need to fork our repository and link it to your RStudio.  

Once you have the project open the first thing you will do is change "Student Name" on line 3 with your name. 
Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A04_Sp21.Rmd"). Submit this pdf using Sakai. 

## Questions 

Consider the same data you used for A2 from the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx". The data comes from the US Energy Information and Administration and corresponds to the January 2021 Monthly Energy Review.

R packages needed for this assignment:"forecast","tseries", and "Kendall". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
#install.packages("readxl")
library("readxl")
library(lubridate)
library(ggplot2)
#install.packages("forecast")
library(forecast)
#install.packages("tseries")
library(tseries)
library(Kendall)

#install.packages("outliers")
library(outliers)
#install.packages("tidyverse")
library(tidyverse)


```


## Stochastic Trend and Stationarity Test

For this part you will once again work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series and the Date column. Don't forget to format the date object.

```{r}
#Importing data set

MonthlyData <- read_excel("../Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",
                          sheet = 1, skip = 9) 

# number of obs
nobsv <- nrow(MonthlyData) 

  
# Select columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption
MonthlyData_subset<- MonthlyData[2:nobsv, c(1, 4, 5, 6)]

# Checking data
head(MonthlyData_subset)
str(MonthlyData_subset)

# Change tbl_df to data frame
MonthlyData_subset = as.data.frame(MonthlyData_subset)
str(MonthlyData_subset)

# number of col
ncoln<- ncol(MonthlyData_subset)-1

# Change column names 
#colnames(MonthlyData_subset)[1] <- "Date"
colnames(MonthlyData_subset)=c("Date","Biomass","Renewable","Hydroelectric")
str(MonthlyData_subset)

# change character format to numeric format
MonthlyData_subset[,2:4] <- sapply(MonthlyData_subset[,2:4],as.numeric)
str(MonthlyData_subset)

# Create a data frame structure with these three time series
# From Jan 1973 to Oct 2020 as a time series object
MonthlyData_subset_ts <- ts(MonthlyData_subset[,2:4], frequency = 12, 
                            start = c(1973, 1, 1), end = c(2020, 10, 1)) 
str(MonthlyData_subset_ts)
```
```{r}
MyDate <- as.Date(MonthlyData_subset$Date)

#create new df
MonthlyData_new <- cbind.data.frame(MyDate, MonthlyData_subset_ts)
str(MonthlyData_new)
class(MonthlyData_new)
head(MonthlyData_new)
```

### Q1
Now let's try to difference these three series using function diff(). Start with the original data from part (b). Try differencing first at lag 1 and plot the remaining series. Did anything change? Do the series still seem to have trend?
```{r}
MonthlyData_subset_ts_diff <- diff(MonthlyData_subset_ts, lag = 1, differences = 1)

# Change MonthlyData_subset_ts_diff to data frame
#df_diff<- as.data.frame(MonthlyData_subset_ts_diff)

#plot
plot(MonthlyData_subset_ts)
plot(MonthlyData_subset_ts_diff)

```

### Q2
Compute Mann-Kendall and Spearman's Correlation Rank Test for each time series. Ask R to print the results. Interpret the results. 

```{r}
#Mann-Kendall Correlation 
#There is seasonality in Hydro data, so use "SeasonalMannKendall"
for(ColNum in 1:3){
  SMKtest_MonthlyData <- SeasonalMannKendall(MonthlyData_subset_ts[,ColNum])
  print("Results for Seasonal Mann Kendall")
  print(summary(SMKtest_MonthlyData))
}
```
>All p value are less than 0.05, so they all have significant correlation between two variables (time and energy data). Biomass and Renweable data has the positive correlation, but hydro one has negative correlation.

```{r}
#Because we know hydro data has seasonality from Assignment3 Question6, 
#so we need to deseason the hydro data first.

EnergyType=3

#First create the seasonal dummies
Energy_dummies <- seasonaldummy(MonthlyData_subset_ts[,EnergyType])

#this function only accepts ts object, no need to add one here because date 
#object is not a column

#Then fit a linear model to the seasonal dummies
seasonal_model=lm(MonthlyData_subset[,(EnergyType+1)]~Energy_dummies)
print(summary(seasonal_model))

#Look at the regression coefficient. These will be the values of Beta

#Store regression coefficients
beta_int=seasonal_model$coefficients[1]
beta_coeff=seasonal_model$coefficients[2:12]

#compute seasonal component
seasonal_comp=array(0,nobsv-1)
for(n in 1:nobsv-1){
  seasonal_comp[n]=(beta_int+beta_coeff%*%Energy_dummies[n,])
}

#Removing seasonal component
Deseason_Hydro_data <- MonthlyData_subset[,(1+EnergyType)]-seasonal_comp

str(Deseason_Hydro_data)
```
```{r}
#change deseason hydro num to data frame
Deseason_Hydro <- as.data.frame(Deseason_Hydro_data)

#create new data frame with Original Biomass, Renewable, and deseason hydro data
df_MonthlyData_new <- 
  subset(MonthlyData_subset, select = -Hydroelectric) %>% 
  cbind(Deseason_Hydro)
head(df_MonthlyData_new)

#Change new data frame to ts format
df_MonthlyData_new_ts <- ts(df_MonthlyData_new[,2:4], frequency = 12, 
                            start = c(1973, 1, 1), end = c(2020, 10, 1)) 
str(df_MonthlyData_new_ts)

```

```{r}
for(ColNum in 1:3){
  #Group data in yearly steps instances
  MonthlyData_new_matrix <- matrix(df_MonthlyData_new_ts[,1],byrow=FALSE,nrow=12)
  YearlyData_mean <- colMeans(MonthlyData_new_matrix)
  
  #library(dplyr)  #move this to package chunk later
  Year <- c(year(first(df_MonthlyData_new$Date)):year(last(df_MonthlyData_new$Date)))
  
  YearlyData <- data.frame(Year, YearlyData_mean)
  str(YearlyData)
  
  print("Results from Spearman Correlation")
  SpCor_MonthlyData=cor.test(YearlyData_mean,Year,method="spearman")
  print(SpCor_MonthlyData)
}

#cor(YearlyData,Year,method="spearman")
```
> All rho are almost 0.9, which shows  strong positive relationship between the Energy data and time. And all p value are less than 0.05, so they all have significant correlation with each other.


## Decomposing the series

For this part you will work only with the following columns: Solar Energy Consumption and Wind Energy Consumption. 

### Q3

Create a data frame structure with these two time series only and the Date column. Drop the rows with *Not Available* and convert the columns to numeric. You can use filtering to eliminate the initial rows or conver to numeric and then use the drop_na() function. If you are familiar with pipes for data wrangling, try using it!

```{r}
#Importing data set
MonthlyData2 <- read_excel("../Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",
                           sheet = 1, skip = 9) 
# number of obs
nobsv <- nrow(MonthlyData2) 

# Select columns: Solar Energy Consumption and Wind Energy Consumption
MonthlyData_subset2<- MonthlyData2[2:nobsv, c(1, 8, 9)] 

# change'Not Available' to "NA"
MonthlyData_subset2[] <- lapply(MonthlyData_subset2, gsub, pattern='Not Available', replacement='NA')

# Checking data
str(MonthlyData_subset2)

# Change column names 
#colnames(MonthlyData_subset2)[1] <- "Date"
colnames(MonthlyData_subset2) <- c("Date","Solar","Wind")
MonthlyData_subset2$Date <- as.Date(MonthlyData_subset2$Date)
str(MonthlyData_subset2)

# change character format to numeric format
MonthlyData_subset2[,2:3] <- sapply(MonthlyData_subset2[,2:3],as.numeric)

# Drop NA
#library(dplyr)
MonthlyData_subset_new <- na.omit(MonthlyData_subset2)
str(MonthlyData_subset_new)
head(MonthlyData_subset_new)
tail(MonthlyData_subset_new)


```

```{r}
# Create a data frame structure with these three time series
# From Jan 1984 as a time series object
MonthlyData_subset_ts2 <- ts(MonthlyData_subset_new[,2:3], frequency = 12, 
                             start = c(1984,1,1),end = c(2020,10,1)) 

MyDate2 <- as.Date(MonthlyData_subset_new$Date)

#create new df
MonthlyData_new2 <- cbind.data.frame(MyDate2, MonthlyData_subset_ts2)
str(MonthlyData_new2)
head(MonthlyData_new2)
```


### Q4

Plot the Solar and Wind energy consumption over time using ggplot. Explore the function scale_x_date() on ggplot and see if you can change the x axis to improve your plot.
Hint: use *scale_x_date(date_breaks = "5 years", date_labels = "%Y")")*

Try changing the color of the wind series to blue. Hint: use *color = "blue"*

```{r}

ggplot(MonthlyData_new2) +
  geom_line(aes(MyDate2, y=Solar, colour = "Solar")) + 
  geom_line(aes(MyDate2, y = Wind, colour = "Wind")) +
  ggtitle("Solar and Wind energy consumption over time") +
  xlab("Year") + 
  ylab("Energy consumption (Trillion Btu)") +
  scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
  scale_colour_manual("", values = c("Solar" = "red", "Wind"= "blue")) +
  theme(axis.text.x = element_text(angle = 45,  hjust = 1),
        plot.title = element_text(hjust=0.5))

```

### Q5

Transform wind and solar series into a time series object and apply the decompose function on them using the additive option. What can you say about the trend component? What about the random component? Does the random component look random? Or does it appear to still have some seasonality on it?

```{r}
str(MonthlyData_subset_ts2)

#Using R decompose function
decompose_Solar_ts2 <- decompose(MonthlyData_subset_ts2[,"Solar"],"additive")
plot(decompose_Solar_ts2)

decompose_Wind_ts2 <- decompose(MonthlyData_subset_ts2[,"Wind"],"additive")
plot(decompose_Wind_ts2)

```
>The trend component of both of them is an increasing pattern. The random component is not that kind of random and it is a regularly repeating pattern before 2015 in Solar dataset and before 2010 in Wind dataset. So there still are some seasonality on that.

### Q6

Use the decompose function again but now change the type of the seasonal component from additive to multiplicative.
What happened to the random component this time? 
```{r}
decompose_Solar_ts2 <- decompose(MonthlyData_subset_ts2[,"Solar"],"multiplicative")
plot(decompose_Solar_ts2)

#par(cex.lab=1.2)
decompose_Wind_ts2 <- decompose(MonthlyData_subset_ts2[,"Wind"],"multiplicative")
plot(decompose_Wind_ts2)

```
>The random outputs are different from "additive" ones. The random pattern occur bedore 1990 for Solar, and occur before 2000 in Wind.

### Q7

When fitting a model to this data, do you think you need all the historical data? Think about the date from 90s and early 20s. Are there any information from those year we might need to forecast the next six months of Solar and/or Wind consumption. Explain your response.
>If there is a consistent pattern for all historical data, then it is useful to fit the model and predict the future trend.
>For Solar and/or Wind consumption dataset, I think the data after 1990 in Solar and after 2000 in Wind should be use to forecast the next six month comsumption.
