---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2021"
author: "Xueying Feng"
subtitle: Assignment 3 - Due date 02/15/21
output:
  pdf_document:
    latex_engine: xelatex
geometry: margin=2.54cm
editor_options:
  chunk_output_type: inline
---
  
## Directions
  
You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. 

Once you have the project open the first thing you will do is change "Student Name" on line 3 with your name. 
Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

Please keep this R code chunk options for the report. It is easier for us to grade when we can see code and output together. And the tidy.opts will make sure that line breaks on your code chunks are automatically added for better visualization. 

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A01_Sp21.Rmd"). Submit this pdf using Sakai. 

## Questions 

Consider the same data you used for A2 from the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx". The data comes from the US Energy Information and Administration and corresponds to the January 2021 Monthly Energy Review. Once again you will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. 

R packages needed for this assignment:"forecast","tseries", and "Kendall". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here

#install.packages("readxl")
library("readxl")
library(lubridate)
library(ggplot2)
#install.packages("forecast")
library(forecast)
#install.packages("tseries")
library(tseries)
library(Kendall)
```

##Trend Component

### Q1
Create a plot window that has one row and three columns. And then for each object on your data frame, fill the plot window with time series plot, ACF and PACF. You may use the some code form A2, but I want all three plots on the same window this time. (Hint: watch videos for M4)
```{r Importing data set}
#Importing data set
MonthlyData <- read_excel("../Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",sheet = 1, skip = 9) 

# number of obs
nobsv <- nrow(MonthlyData) 

  
# Select columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption
MonthlyData_subset<- MonthlyData[2:nobsv, c(1, 4, 5, 6)]

# Checking data
head(MonthlyData_subset)
str(MonthlyData_subset)

# Change tbl_df to data frame
MonthlyData_subset = as.data.frame(MonthlyData_subset)
str(MonthlyData_subset)
ncoln<- ncol(MonthlyData_subset)-1
```

```{r Transforming data into time series object}

# change character format to numeric format
MonthlyData_subset[,2:4] <- sapply(MonthlyData_subset[,2:4],as.numeric)
str(MonthlyData_subset)

# Change column names 
colnames(MonthlyData_subset)[1] <- "Date"
#colnames(MonthlyData_subset)=c("Date","Biomass","Renewable","Hydroelectric")
str(MonthlyData_subset)
```

```{r plot}
#using package ggplot2
#devtools::install_github('cran/ggplot2')


for(i in 1:ncoln){
  par(mfrow=c(1,3))
  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+i)])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(MonthlyData_subset)[(1+i)]," (Trillion Btu)",sep="")) 
        )
}
```

```{r Time series}
# Create a data frame structure with these three time series
# From Jan 1973 to Oct 2020 as a time series object
MonthlyData_subset_ts <- ts(MonthlyData_subset[,2:4], frequency = 12, start = c(1973, 1, 1), end = c(2020, 10, 1)) 
str(MonthlyData_subset_ts)
head(MonthlyData_subset_ts)
tail(MonthlyData_subset_ts)
```

```{r plot: ACF and PACF}
# number of obs
#ncoln<- ncol(MonthlyData_subset)

#Acf and Pacf
par(mfrow=c(1,3))  #place plot side by side
for(i in 1:ncoln){
  # because I am not storing Acf() into any object, I don't need to specify plot=TRUE 
  Acf(MonthlyData_subset_ts[,i],lag.max=40,main=paste("AFC of ",colnames(MonthlyData_subset)[(1+i)],sep="")) 
}

par(mfrow=c(1,3))  #place plot side by side
for(i in 1:ncoln){
  Pacf(MonthlyData_subset_ts[,i],lag.max=40,main=paste("PAFC of ",colnames(MonthlyData_subset)[(1+i)],sep=""))
}

```

### Q2
From the plot in Q1, do the series Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption appear to have a trend? If yes, what kind of trend?

>The series Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption appear to have a trend. 
>Total Biomass Energy Production showsa clear upward trend overall, but 1990-2000 shows kind of random variation. >Total Renewable Energy Production: before 1980 and after 2002 show a clear upward trend, between 1985 and 2002, the cycles do not repeat at regular intervals and do not have the same shape. 
>Hydroelectric Power Consumption shows cyclic movements, but in general is a decreasing trend.


### Q3
Use the *lm()* function to fit a linear trend to the three time series. Ask R to print the summary of the regression. Interpret the regression output, i.e., slope and intercept. Save the regression coefficients for further analysis.
```{r lm()}
nobsv <-  nrow(MonthlyData_subset) #int
nobsv <- nobsv-1+1 #numeric

#Create vector n
n <- c(1:nobsv)

par(mfrow=c(1,3))
for(EnergyType in 1:ncoln){
  #Fit a linear trend to TS of EnergyType
  LinearTrend_model <- lm(MonthlyData_subset[,EnergyType+1]~n)  
  print(summary(LinearTrend_model))
  
  beta0=as.numeric(LinearTrend_model$coefficients[1])  #first coefficient is the intercept term or beta0
  beta1=as.numeric(LinearTrend_model$coefficients[2])  #second coefficient is the slope or beta1
  print(beta0)
  print(beta1)
  
  #Let's plot the time series with its trend line （Command+Shift+C）
  # print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
  #             geom_line(color="blue") +
  #             ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
  #             #geom_abline(intercept = beta0, slope = beta1, color="red")
  #             geom_smooth(color="red",method="lm") )
}

```
>Formula Call:the output is the formula R used to fit the data. The formula just needs the predictor (vector n) and the target/response variable (production), together with the data being used (MonthlyData_subset).
>The Residuals section of the model output breaks it down into 5 summary points. When assessing how well the model fit the data, you should look for a symmetrical distribution across these points on the mean value zero (0).
>In simple linear regression, the coefficients are two unknown constants that represent the intercept and slope terms in the linear model. 
>The coefficient Standard Error measures the average amount that the coefficient estimates vary from the actual average value of our response variable.
>A small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between predictor and response variable.
>The R-squared statistic provides a measure of how well the model is fitting the actual data. 
 
### Q4
Use the regression coefficients from Q3 to detrend the series. Plot the detrended series and compare with the plots from Q1. What happened? Did anything change?
```{r Plot the detrended series}
#remove the trend from series
par(mfrow=c(1,3))

for(EnergyType in 1:ncoln){
  #Fit a linear trend to TS of EnergyType
  LinearTrend_model <- lm(MonthlyData_subset[,EnergyType+1]~n)  
  summary(LinearTrend_model)
  
  beta0=as.numeric(LinearTrend_model$coefficients[1])  #first coefficient is the intercept term or beta0
  beta1=as.numeric(LinearTrend_model$coefficients[2])  #second coefficient is the slope or beta1
  
  #Remove the trend
  DetrendInflow_data <- MonthlyData_subset[,(EnergyType+1)]-(beta0+beta1*n)

  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
              geom_line(color="blue") +
              ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
              #geom_abline(intercept = beta0, slope = beta1, color="red")
              geom_smooth(color="red",method="lm") +
              geom_line(aes(y=DetrendInflow_data), col="green")+
              geom_smooth(aes(y=DetrendInflow_data),color="orange",method="lm"))
}
```
>First, Detrend plot makes a time series more smoothed and stationary, which does not have obvious increasing or decreasing trend. Second, the value changes near the 0.


### Q5
Plot ACF and PACF for the detrended series and compare with the plots from Q1. Did the plots change? How?
```{r Detrend data frame}
Data_Detrend <- MonthlyData_subset

# Add detrend data column in data frame
for(i in 1:ncoln) {
  # Create new column
  detrend <- rep(i, nrow(Data_Detrend))
  # Append new column
  Data_Detrend[ , ncol(Data_Detrend) + 1] <- MonthlyData_subset[,(i+1)]-(beta0+beta1*n)
  # Rename column name
  colnames(Data_Detrend)[ncol(Data_Detrend)] <- paste0(colnames(Data_Detrend)[(1+i)]," Detrend",sep="")  
}

# Check data
str(Data_Detrend)
head(Data_Detrend)
```

```{r Detrend Acf and Pacf plots}
#change Detrend_data to TS
Data_Detrend_ts <- ts(Data_Detrend[,2:7],frequency=12) 
str(Data_Detrend_ts)

#Acf and Pacf plots
for(i in 1:ncoln){
  par(mfrow=c(1,2))  #place plot side by side
  Acf(Data_Detrend_ts[,i+3],lag.max=40,main=paste("AFC of ",colnames(Data_Detrend_ts)[(3+i)],sep="")) 
  # because I am not storing Acf() into any object, I don't need to specify plot=TRUE 
  Pacf(Data_Detrend_ts[,i+3],lag.max=40,main=paste("PAFC of ",colnames(Data_Detrend_ts)[(3+i)],sep=""))
}
```
>The plot change a little bit, but not too much for plot 1 and 2. 
>The third ACF at lag 5 6 7 8 are changed to nagetive and less lines fall into two dotted lines.

## Seasonal Component

Set aside the detrended series and consider the original series again from Q1 to answer Q6 to Q8.

### Q6
Do the series seem to have a seasonal trend? Which serie/series? Use function *lm()* to fit a seasonal means model to this/these time series. Ask R to print the summary of the regression. Interpret the regression output. Save the regression coefficients for further analysis.
```{r Seasonal trend}

for(EnergyType in 1:ncoln){
  #Use seasonal means model
  #First create the seasonal dummies
  Energy_dummies <- seasonaldummy(MonthlyData_subset_ts[,EnergyType])  
  #this function only accepts ts object, no need to add one here because date 
  #object is not a column
  
  #Then fit a linear model to the seasonal dummies
  seasonal_model=lm(MonthlyData_subset[,(EnergyType+1)]~Energy_dummies)
  print(summary(seasonal_model))
  
  #Look at the regression coefficient. These will be the values of Beta
  
  #Store regression coefficients
  beta_int=seasonal_model$coefficients[1]
  beta_coeff=seasonal_model$coefficients[2:12]
  
  #compute seasonal component
  seasonal_comp=array(0,nobsv)
  for(n in 1:nobsv){
    seasonal_comp[n]=(beta_int+beta_coeff%*%Energy_dummies[n,])
  }
  
  #Understanding what we did
  # print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
  #             geom_line(color="blue") +
  #             ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
  #             geom_line(aes(y=seasonal_comp), col="red"))
}
```
>There series all seem to have a seasonal trend.


### Q7
Use the regression coefficients from Q6 to deseason the series. Plot the deseason series and compare with the plots from part Q1. Did anything change?
```{r}
for(EnergyType in 1:ncoln){
  
  #Removing seasonal component
  Deseason_data <- MonthlyData_subset[,(1+EnergyType)]-seasonal_comp
  
  #Understanding what we did
  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
              geom_line(color="blue") +
              ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
              geom_line(aes(y=Deseason_data), col="green"))
}
```
>First plot: the magnitude of the change increases compared to original one, which means has more obvious oscillations. The other two plots has less obvious oscillations.Moreover, the values are smaller than original ones.



### Q8
Plot ACF and PACF for the deseason series and compare with the plots from Q1. Did the plots change? How?
```{r ACF and PACF for the deseason series}

for(EnergyType in 1:ncoln){
  Deseason_data <- MonthlyData_subset[,(1+EnergyType)]-seasonal_comp
  str(Deseason_data)
  
  #change Deseason_data to TS
  Deseason_data_ts <- ts(Deseason_data, frequency = 12) 
  
  #plot ACF and PACF for Detrend_data_ts
  par(mfrow=c(1,2)) 
  Acf(Deseason_data_ts,lag.max=40,main=paste("AFC of ",colnames(MonthlyData_subset)[(1+i)],sep="")) 
  Pacf(Deseason_data_ts,lag.max=40,main=paste("PAFC of ",colnames(MonthlyData_subset)[(1+i)],sep=""))
}
  
```
> The spike values of ACF in plot 1 and 2 are different from that of plots from Q1. Plot 1 has more obvious changes, but plot 2 is less obvious.
> For Thrid plot, the value of ACF are all positive, which is total different from Q1.
