---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2021"
author: "Xueying Feng"
subtitle: Assignment 6 - Due date 03/26/21
output:
  pdf_document: default
  html_document:
    df_print: paged
geometry: margin=2.54cm
header-includes:
- \usepackage{enumerate}
- \usepackage{enumitem}
editor_options:
  chunk_output_type: inline
---

## Directions

You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. And to do so you will need to fork our repository and link it to your RStudio.

Once you have the project open the first thing you will do is change "Student Name" on line 3 with your name. Then you will start working through the assignment by **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A06_Sp21.Rmd"). Submit this pdf using Sakai.

## Set up

```{r}
#Load/install required package here
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)

#install.packages("kableExtra")
library(kableExtra)

```

## Importing and processing the data set

Consider the data from the file "Net_generation_United_States_all_sectors_monthly.csv". The data corresponds to the monthly net generation from January 2001 to December 2020 by source and is provided by the US Energy Information and Administration. **You will work with the natural gas column only**.

Packages needed for this assignment: "forecast","tseries". Do not forget to load them before running your script, since they are NOT default packages.\\

### Q1

Import the csv file and create a time series object for natural gas. Make you sure you specify the **start=** and **frequency=** arguments. Plot the time series over time, ACF and PACF.

```{r}
NetGen <- read.csv("../Data/Net_generation_United_States_all_sectors_monthly.csv",skip = 4)
head(NetGen)


#Inspect data
head(NetGen)
ncol <- ncol(NetGen)
nobs <- nrow(NetGen)

# Change column names 
#colnames(NetGen)[1] <- "Date"
colnames(NetGen)=c("Date","AllFuels","Coal","NG", "Nuclear","ConventionalHydroelectric")
str(NetGen)
head(NetGen)

#convert to numeric 
#NetGen[,2:ncol] <- sapply(NetGen[,2:ncol],as.numeric)

# change character format to numeric format
#library(lubridate)
NetGen$Date <- parse_date_time(NetGen$Date,"by")

#arrange date from 2001 to 2020
NetGen<- NetGen %>% 
  arrange(NetGen$Date)

str(NetGen)
head(NetGen)

#Transforming data into time series object
ts_NetGen <- ts(NetGen[,2:(ncol)],
                start=c(year(NetGen$Date[1]),
                        month(NetGen$Date[1])),
                frequency=12)
str(ts_NetGen)


#plot NG trend
plot(ts_NetGen[,3])


#ACF and PACF plots
par(mar=c(3,3,3,0));par(mfrow=c(1,2))
ACF_Plot <- Acf(NetGen$NG, lag = 40, plot = TRUE)
  ## ACF should seasonality, and has correlation, it is not stationary 
PACF_Plot <- Pacf(NetGen$NG, lag = 40)
  ## Pacf plot shows we might dealing with stochastic or unit root (means random)


```

### Q2

Using the $decompose()$ or $stl()$ and the $seasadj()$ functions create a series without the seasonal component, i.e., a deseasonalized natural gas series. Plot the deseasonalized series over time and corresponding ACF and PACF. Compare with the plots obtained in Q1.

```{r}
#Using R decompose function
decompose_ts_NetGen <- decompose(ts_NetGen[,"NG"],"additive")
plot(decompose_ts_NetGen)
#decompose_ts_NetGen2 <- decompose(ts_NetGen[,"NG"],"multiplicative")
#plot(decompose_ts_NetGen2)

##Using R seasadj function
seasadj_ts_NetGen <- seasadj(decompose_ts_NetGen) 
plot(seasadj_ts_NetGen)

par(mar=c(3,3,3,0));par(mfrow=c(1,2))
seasadj_ACF_Plot <- Acf(seasadj_ts_NetGen, lag = 40, plot = TRUE)
seasadj_PACF_Plot <- Pacf(seasadj_ts_NetGen, lag = 40)
```

## Modeling the seasonally adjusted or deseasonalized series

### Q3

Run the ADF test and Mann Kendall test on the deseasonalized data from Q2. Report and explain the results.

```{r ADF}
#Run ADF
ADFtest <- adf.test(seasadj_ts_NetGen,alternative="stationary")
print(ADFtest)

#Run MannKendall
library(Kendall)
print("Results of Mann Kendall on average yearly series")
print(summary(MannKendall(seasadj_ts_NetGen)))
```
>Augmented Dickey-Fuller test:  the p-value of the test is lower than significance level 0.05, then it is stationary.
>Mann Kendall: the p-value of the test is lower than  significance level 0.05 , then there is statistically significant evidence that a trend is present in the time series data.


### Q4

Using the plots from Q2 and test results from Q3 identify the ARIMA model parameters $p,d$ and $q$. Note that in this case because you removed the seasonal component prior to identifying the model you don't need to worry about seasonal component. Clearly state your criteria and any additional function in R you might use. DO NOT use the $auto.arima()$ function. You will be evaluated on ability to can read the plots and interpret the test results.

>The stationary series has positive autocorrelation at lag 1, then autoregressive model(AR) terms work best. 
It is a autoregressive model with p=2 also because of a slow decay in the ACF plot and a clear cut off at lag 2 in the PACF plot (p=1).
We assume we are  working with a zero-mean process,leading to the conclusion that d=0.


### Q5

Use $Arima()$ from package "forecast" to fit an ARIMA model to your series considering the order estimated in Q4. Should you allow for constants in the model, i.e., $include.mean = TRUE$ or $include.drift=TRUE$. **Print the coefficients** in your report. Hint: use the $cat()$ function to print.

```{r}
Arima.model <- Arima(seasadj_ts_NetGen,order=c(1,0,0),include.mean=TRUE)
print(Arima.model)

#cat("coefficients is:", )                       


```

### Q6

Now plot the residuals of the ARIMA fit from Q5 along with residuals ACF and PACF on the same window. You may use the $checkresiduals()$ function to automatically generate the three plots. Do the residual series look like a white noise series? Why?
```{r}
checkresiduals(Arima.model)
#The residual series do not look like a white noise series, because it is not between two blue dotted lines.

```

## Modeling the original series (with seasonality)

### Q7

Repeat Q4-Q6 for the original series (the complete series that has the seasonal component). Note that when you model the seasonal series, you need to specify the seasonal part of the ARIMA model as well, i.e., $P$, $D$ and $Q$.

```{r}
#ACF and PACF plots
par(mar=c(3,3,3,0));par(mfrow=c(1,2))
ACF_Plot <- Acf(NetGen$NG, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(NetGen$NG, lag = 40)

#For the order of the non-seasonal component, we focus on the early lags (ex.before lag=12). We can see that ACF shows a slow decay and PACF has a clear cut off at lag 2, which indicates that the non-seasonal component has AR process, p=2.
#As for the seasonal component, we are only intrested at seasonal lags 12, 24, 36 and so forth. We can see that the ACF has multiple spikes at the seasonal lags and PACF has only one spike at lag 12. It shows that we should include a SAR process, P=1.

Arima.Sea.odel <-Arima(ts_NetGen[,"NG"], order=c(1,0,0), seasonal=list(order=c(1,0,0),include.mean=TRUE, period=12))
summary(Arima.Sea.odel)
checkresiduals(Arima.Sea.odel)
#The residual series do look like a white noise series, because only one spike is not in two dotted line. 

```


### Q8

Compare the residual series for Q7 and Q6. Can you tell which ARIMA model is better representing the Natural Gas Series? Is that a fair comparison? Explain your response.

```{r}
#I can tell which ARIMA model is better representing the Natural Gas Series, and second one is better.But I dont think that is a fair comparison, because the pdq and PDQ value that we tell from plot is not accurate.
```


## Checking your model with the auto.arima()

**Please** do not change your answers for Q4 and Q7 after you ran the $auto.arima()$. It is **ok** if you didn't get all orders correctly. You will not loose points for not having the correct orders. The intention of the assignment is to walk you to the process and help you figure out what you did wrong (if you did anything wrong!).

### Q9

Use the $auto.arima()$ command on the **deseasonalized series** to let R choose the model parameter for you. What's the order of the best ARIMA model? Does it match what you specified in Q4?

```{r}
SARIMA_autofit <- auto.arima(seasadj_ts_NetGen)
print(SARIMA_autofit)
checkresiduals(SARIMA_autofit)
# order is 1, and it only matches the p value what I specified in Q4
```


### Q10

Use the $auto.arima()$ command on the **original series** to let R choose the model parameters for you. Does it match what you specified in Q7?

```{r}
SARIMA_autofit <- auto.arima(ts_NetGen[,"NG"])
checkresiduals(SARIMA_autofit)
#It does not matches what I specified in Q4.

```

