---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2021"
subtitle: "Assignment 2 - Due date 02/05/21"
author: Xueying Feng
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## Submission Instructions

You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github.   

Once you have the file open on your local machine the first thing you will do is change "Student Name" on line 4 with your name. 
Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A02_Sp21.Rmd"). Submit this pdf using Sakai. 

## R packages

R packages needed for this assignment:"forecast","tseries", and "dplyr". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
getwd()

library(lubridate)
library(ggplot2)

library(dplyr)
#install.packages("forecast")
library(forecast)
#install.packages("tseries")
library(tseries)
```


## Data set information  

Consider the data provided in the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx" on our **Data** folder. The data comes from the US Energy Information and Administration and corresponds to the January 2021 Monthly Energy Review. The spreadsheet is ready to be used. Use the command $read.table()$ to import the data in R or $panda.read\_excel()$ in Python (note that you will need to import pandas package). }


```{r Importing data set}
#Importing data set

#install.packages("readxl")
library("readxl")
MonthlyData <- read_excel("./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",sheet = 1, skip = 9) 
MonthlyData

#AnnualData <- read_excel("./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",2)
#AnnualData
```

## Question 1
You will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. Use the command head() to verify your data.

```{r}
#specific rows and columns just trim the table
MonthlyData_subset<- MonthlyData[2:575, c(1, 4, 5, 6)]
MonthlyData_subset

#Checking data
head(MonthlyData_subset)
str(MonthlyData_subset)
```


## Question 2
Transform your data frame in a time series object and specify the starting point and frequency of the time series using the function ts().
```{r}
# change character format to numeric format
MonthlyData_subset[,2:4] <- sapply(MonthlyData_subset[,2:4],as.numeric)
str(MonthlyData_subset)

#add column names 
colnames(MonthlyData_subset)=c("Month","Biomass","Renewable","Hydroelectric")


# from Jan 1973 to Oct 2020 as a time series object
MonthlyData_subset_ts <- ts(MonthlyData_subset[2:4], frequency = 12, start = c(1973, 1, 1), end = c(2020, 10, 1)) 
MonthlyData_subset_ts
```

## Question 3
Compute mean and standard deviation for these three series.
```{r}
Biomass_mean <- mean(MonthlyData_subset_ts[,"Biomass"])
Biomass_sd <- sd(MonthlyData_subset_ts[,"Biomass"])
Biomass_mean
str(Biomass_mean)
Biomass_sd

Renewable_mean <- mean(MonthlyData_subset_ts[,"Renewable"])
Renewable_sd <- sd(MonthlyData_subset_ts[,"Renewable"])
Renewable_mean
Renewable_sd

Hydroelectric_mean <- mean(MonthlyData_subset_ts[,"Hydroelectric"])
Hydroelectric_sd <- sd(MonthlyData_subset_ts[,"Hydroelectric"])
Hydroelectric_mean
Hydroelectric_sd

```


## Question 4
Display and interpret the time series plot for each of these variables. Try to make your plot as informative as possible by writing titles, labels, etc. For each plot add a horizontal line at the mean of each series in a different color.
```{r}
library(ggplot2)
#plot.new()

#using package ggplot2
Biomass_plot <- ggplot(MonthlyData_subset, aes(x=Month, y=Biomass)) +
  geom_line(aes(y=Biomass,col="Total Biomass Energy Production"),group = 1) +
  geom_line(aes(y=Biomass_mean,col="Biomass mean"),group = 2) +
  ggtitle("Total Biomass Energy Production Data") +
  xlab("Time") + 
  ylab("Total Biomass Energy Production (Trillion Btu)")
Biomass_plot

#dev.off()

```

```{r}
Renewable_plot <- ggplot(MonthlyData_subset, aes(x=Month, y=Renewable)) +
  geom_line(aes(y=Renewable,col="Total Renewable Energy Production"),group = 1) +
  geom_line(aes(y=Renewable_mean,col="Renewable mean"),group = 2) +
  ggtitle("Total Renewable Energy Production Data") +
  xlab("Time") + 
  ylab("Total Renewable Energy Production (Trillion Btu)")
Renewable_plot
```

```{r}
Hydroelectric_plot <- ggplot(MonthlyData_subset, aes(x=Month, y=Hydroelectric)) +
  geom_line(aes(y=Hydroelectric,col="Hydroelectric Power Consumption"),group = 1) +
  geom_line(aes(y=Hydroelectric_mean,col="Hydroelectric mean"),group = 2) +
  ggtitle("Hydroelectric Power Consumption Data") +
  xlab("Time") + 
  ylab("Hydroelectric Power Consumption (Trillion Btu)")
Hydroelectric_plot
```

## Question 5
Compute the correlation between these three series. Are they significantly correlated? Explain your answer.
```{r}

pairs(MonthlyData_subset[, c(2,3,4)])

cor(MonthlyData_subset$Biomass, MonthlyData_subset$Renewable)

cor(MonthlyData_subset$Biomass, MonthlyData_subset$Hydroelectric)

cor(MonthlyData_subset$Renewable, MonthlyData_subset$Hydroelectric)

#Due to the results of correlation and plots, Total Biomass Energy Production and Hydroelectric Power Consumption  are significatly correlated, whereas they other two are not.
```


## Question 6
Compute the autocorrelation function from lag 1 up to lag 40 for these three variables. What can you say about these plots? Do the three of them have the same behavior?

```{r ACF}
Biomass_acf = acf(MonthlyData_subset_ts[,1], lag.max=40, type = "correlation", plot=TRUE, main="Biomass ACF")

Renewable_acf = acf(MonthlyData_subset_ts[,2], lag.max=40, type = "correlation", plot=TRUE, main="Renewable ACF")

Hydroelectric_acf = acf(MonthlyData_subset_ts[,3], lag.max=40, type = "correlation", plot=TRUE, main="Hydroelectric ACF")




```


## Question 7
Compute the partial autocorrelation function from lag 1 to lag 40 for these three variables. How these plots differ from the ones in Q6?
```{r PACF}
Biomass_pacf = pacf(MonthlyData_subset_ts[,1], lag.max=40, type = "correlation", plot=TRUE, main="Biomass PACF")

Renewable_pacf = pacf(MonthlyData_subset_ts[,2], lag.max=40, type = "correlation", plot=TRUE, main="Renewable PACF")

Hydroelectric_acf = acf(MonthlyData_subset_ts[,3], lag.max=40, type = "correlation", plot=TRUE, main="Hydroelectric PACF")



```

