---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2021"
subtitle: "Assignment 3 - Due date 02/12/21"
author: "Xueying Feng"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: inline
---
  
## Directions
  
You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. 

Once you have the project open the first thing you will do is change "Student Name" on line 3 with your name. 
Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

Please keep this R code chunk options for the report. It is easier for us to grade when we can see code and output together. And the tidy.opts will make sure that line breaks on your code chunks are automatically added for better visualization. 

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A01_Sp21.Rmd"). Submit this pdf using Sakai. 

## Questions 

Consider the same data you used for A2 from the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx". The data comes from the US Energy Information and Administration and corresponds to the January 2021 Monthly Energy Review. Once again you will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. 

R packages needed for this assignment:"forecast","tseries", and "Kendall". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here

#install.packages("readxl")
library("readxl")
library(lubridate)
library(ggplot2)
#install.packages("forecast")
library(forecast)
#install.packages("tseries")
library(tseries)
library(Kendall)
```

##Trend Component

### Q1
Create a plot window that has one row and three columns. And then for each object on your data frame, fill the plot window with time series plot, ACF and PACF. You may use the some code form A2, but I want all three plots on the same window this time. (Hint: watch videos for M4)
```{r Importing data set}
#Importing data set
MonthlyData <- read_excel("../Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",sheet = 1, skip = 9) 

# number of obs
nobsv <- nrow(MonthlyData) 

  
# Select columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption
MonthlyData_subset<- MonthlyData[2:nobsv, c(1, 4, 5, 6)]

# Checking data
head(MonthlyData_subset)
str(MonthlyData_subset)

# Change tbl_df to data frame
MonthlyData_subset = as.data.frame(MonthlyData_subset)
str(MonthlyData_subset)
ncoln<- ncol(MonthlyData_subset)-1
```

```{r Transforming data into time series object}

# change character format to numeric format
MonthlyData_subset[,2:4] <- sapply(MonthlyData_subset[,2:4],as.numeric)
str(MonthlyData_subset)

# Change column names 
colnames(MonthlyData_subset)[1] <- "Date"
#colnames(MonthlyData_subset)=c("Date","Biomass","Renewable","Hydroelectric")
str(MonthlyData_subset)
```

```{r plot}
#using package ggplot2
#devtools::install_github('cran/ggplot2')

for(i in 1:ncoln){
  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+i)])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(MonthlyData_subset)[(1+i)]," (Trillion Btu)",sep="")) 
        )
}
```

```{r Time series}
# Create a data frame structure with these three time series
# From Jan 1973 to Oct 2020 as a time series object
MonthlyData_subset_ts <- ts(MonthlyData_subset[,2:4], frequency = 12, start = c(1973, 1, 1), end = c(2020, 10, 1)) 
str(MonthlyData_subset_ts)
head(MonthlyData_subset_ts)
tail(MonthlyData_subset_ts)
```

```{r plot: ACF and PACF}
# number of obs
#ncoln<- ncol(MonthlyData_subset)

#Acf and Pacf
par(mfrow=c(1,3))  #place plot side by side
for(i in 1:ncoln){
  # because I am not storing Acf() into any object, I don't need to specify plot=TRUE 
  Acf(MonthlyData_subset_ts[,i],lag.max=40,main=paste("AFC of ",colnames(MonthlyData_subset)[(1+i)],sep="")) 
}

par(mfrow=c(1,3))  #place plot side by side
for(i in 1:ncoln){
  Pacf(MonthlyData_subset_ts[,i],lag.max=40,main=paste("PAFC of ",colnames(MonthlyData_subset)[(1+i)],sep=""))
}

```

### Q2
From the plot in Q1, do the series Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption appear to have a trend? If yes, what kind of trend?
```{r}
# The Biomass and Renewable plots have very similar trend, but Hydroelectric one is kind of a periodic trend. They do not have same behavior.

# The three PACF plots are much more similar.
```

### Q3
Use the *lm()* function to fit a linear trend to the three time series. Ask R to print the summary of the regression. Interpret the regression output, i.e., slope and intercept. Save the regression coefficients for further analysis.
```{r lm()}
nobsv <-  nrow(MonthlyData_subset) #int
nobsv <- nobsv-1+1 #numeric

#Create vector n
n <- c(1:nobsv)

par(mfrow=c(1,3))
for(EnergyType in 1:ncoln){
  #Fit a linear trend to TS of EnergyType
  LinearTrend_model <- lm(MonthlyData_subset[,EnergyType+1]~n)  
  summary(LinearTrend_model)
  
  beta0=as.numeric(LinearTrend_model$coefficients[1])  #first coefficient is the intercept term or beta0
  beta1=as.numeric(LinearTrend_model$coefficients[2])  #second coefficient is the slope or beta1
  
  #Let's plot the time series with its trend line
  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
              geom_line(color="blue") +
              ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
              #geom_abline(intercept = beta0, slope = beta1, color="red")
              geom_smooth(color="red",method="lm") )
}

```

 
### Q4
Use the regression coefficients from Q3 to detrend the series. Plot the detrended series and compare with the plots from Q1. What happened? Did anything change?
```{r Plot the detrended series}
#remove the trend from series
par(mfrow=c(1,3))

for(EnergyType in 1:ncoln){
  #Fit a linear trend to TS of EnergyType
  LinearTrend_model <- lm(MonthlyData_subset[,EnergyType+1]~n)  
  summary(LinearTrend_model)
  
  beta0=as.numeric(LinearTrend_model$coefficients[1])  #first coefficient is the intercept term or beta0
  beta1=as.numeric(LinearTrend_model$coefficients[2])  #second coefficient is the slope or beta1
  
  #Remove the trend
  DetrendInflow_data <- MonthlyData_subset[,(EnergyType+1)]-(beta0+beta1*n)

  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
              geom_line(color="blue") +
              ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
              #geom_abline(intercept = beta0, slope = beta1, color="red")
              geom_smooth(color="red",method="lm") +
              geom_line(aes(y=DetrendInflow_data), col="green")+
              geom_smooth(aes(y=DetrendInflow_data),color="orange",method="lm"))
}
```


### Q5
Plot ACF and PACF for the detrended series and compare with the plots from Q1. Did the plots change? How?
```{r Detrend data frame}
Data_Detrend <- MonthlyData_subset

# Add detrend data column in data frame
for(i in 1:ncoln) {
  # Create new column
  detrend <- rep(i, nrow(Data_Detrend))
  # Append new column
  Data_Detrend[ , ncol(Data_Detrend) + 1] <- MonthlyData_subset[,(i+1)]-(beta0+beta1*n)
  # Rename column name
  colnames(Data_Detrend)[ncol(Data_Detrend)] <- paste0(colnames(Data_Detrend)[(1+i)]," Detrend",sep="")  
}

# Check data
str(Data_Detrend)
head(Data_Detrend)
```
```{r Detrend Acf and Pacf plots}
#change Detrend_data to TS
Data_Detrend_ts <- ts(Data_Detrend[,2:7],frequency=12) 
str(Data_Detrend_ts)

#Acf and Pacf plots
for(i in 1:ncoln){
  par(mfrow=c(1,2))  #place plot side by side
  Acf(Data_Detrend_ts[,i+3],lag.max=40,main=paste("AFC of ",colnames(Data_Detrend_ts)[(3+i)],sep="")) 
  # because I am not storing Acf() into any object, I don't need to specify plot=TRUE 
  Pacf(Data_Detrend_ts[,i+3],lag.max=40,main=paste("PAFC of ",colnames(Data_Detrend_ts)[(3+i)],sep=""))
}
```

## Seasonal Component

Set aside the detrended series and consider the original series again from Q1 to answer Q6 to Q8.

### Q6
Do the series seem to have a seasonal trend? Which serie/series? Use function *lm()* to fit a seasonal means model to this/these time series. Ask R to print the summary of the regression. Interpret the regression output. Save the regression coefficients for further analysis.
```{r Seasonal trend}

for(EnergyType in 1:ncoln){
  #Use seasonal means model
  #First create the seasonal dummies
  Energy_dummies <- seasonaldummy(MonthlyData_subset_ts[,EnergyType])  
  #this function only accepts ts object, no need to add one here because date 
  #object is not a column
  
  #Then fit a linear model to the seasonal dummies
  seasonal_model=lm(MonthlyData_subset[,(EnergyType+1)]~Energy_dummies)
  summary(seasonal_model)
  
  #Look at the regression coefficient. These will be the values of Beta
  
  #Store regression coefficients
  beta_int=seasonal_model$coefficients[1]
  beta_coeff=seasonal_model$coefficients[2:12]
  
  #compute seasonal component
  seasonal_comp=array(0,nobsv)
  for(n in 1:nobsv){
    seasonal_comp[n]=(beta_int+beta_coeff%*%Energy_dummies[n,])
  }
  
  #Understanding what we did
  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
              geom_line(color="blue") +
              ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
              geom_line(aes(y=seasonal_comp), col="red"))
}
```



### Q7
Use the regression coefficients from Q6 to deseason the series. Plot the deseason series and compare with the plots from part Q1. Did anything change?
```{r}
for(EnergyType in 1:ncoln){
  
  #Removing seasonal component
  Deseason_data <- MonthlyData_subset[,(1+EnergyType)]-seasonal_comp
  
  #Understanding what we did
  print(ggplot(MonthlyData_subset, aes(x=Date, y=MonthlyData_subset[,(1+EnergyType)])) +
              geom_line(color="blue") +
              ylab(paste0(colnames(MonthlyData_subset)[(1+EnergyType)]," (Trillion Btu)",sep="")) +
              geom_line(aes(y=Deseason_data), col="green"))
}
```

### Q8
Plot ACF and PACF for the deseason series and compare with the plots from Q1. Did the plots change? How?
```{r ACF and PACF for the deseason series}

for(EnergyType in 1:ncoln){
  Deseason_data <- MonthlyData_subset[,(1+EnergyType)]-seasonal_comp
  str(Deseason_data)
  
  #change Deseason_data to TS
  Deseason_data_ts <- ts(Deseason_data, frequency = 12) 
  
  #plot ACF and PACF for Detrend_data_ts
  par(mfrow=c(1,2)) 
  Acf(Deseason_data_ts,lag.max=40,main=paste("AFC of ",colnames(MonthlyData_subset)[(1+i)],sep="")) 
  Pacf(Deseason_data_ts,lag.max=40,main=paste("PAFC of ",colnames(MonthlyData_subset)[(1+i)],sep=""))
}
  
```

